---
title: The Tidymodels Workflow
author: Caroline Ledbetter
date: '2019-12-27'
publishDate: '2020-01-31'
slug: tidymodels-suite-demo
categories:
  - R
tags:
  - tidyverse
  - tidymodels
lastmod: '2019-12-27T17:14:17-07:00'
description: ''
show_in_homepage: yes
show_description: 'Using the tidymodels package from beginning to end'
license: ''
featured_image: ''
featured_image_preview: ''
comment: yes
math: no
---

```{r echo=FALSE}
blogdown::shortcode('tweet', '1204918320346157056')
```

From the tidymodels readme: 
I have been trying to incorporate more of the `tidymodels` suite of packages 
into my predictive models workflow (`rsample`, `recipes`) but I find myself 
frequently falling back on `caret` for model tuning and fitting 
because it's what I know. This post is a work through from start to finish using 
the `tidymodels` suite. This post is NOT a tutorial for supervised learning. 
It assumes you know how and why to split your data, resample, up/down sample 
tune parameters, evaluate models etc. 
If you are looking for guides for machine learning, I highly recommend:  
:star: 
[Learning to teach machines to learn:](https://alison.rbind.io/post/2019-12-23-learning-to-teach-machines-to-learn/) 
This post from Alison Hill is full of great resources.  
:closed_book:[Applied Predictive Modeling](http://appliedpredictivemodeling.com) 
by Max Kuhn and Kjell Johnson  
:closed_book:
[Intro to Statisitcal Learning](https://www.springer.com/gp/book/9781461471370)
by Gareth James, Daniela Witten, Trevor Hastie and Robert Tibshirani  
:closed_book:
[The Elements of Statistical Learning](https://www.springer.com/gp/book/9780387848570)
by Trevor Hastie, Robert Tibshirani and Jerome Friedman  

**The Data**  
For this post I am going to use the 
[German Credit data](http://archive.ics.uci.edu/ml/datasets/Statlog+(German+Credit+Data)) 
from the University of California Irving Machine Learning Repository. This 
data set is included in the caret package, but the categorical variables are 
dummy coded and I want to demonstrate some of the factor options in recipe, so 
I am retrieving the original. 

```{r, message = FALSE}
library(tidymodels)
library(readr)
german_credit <- 
  read_table2("http://archive.ics.uci.edu/ml/machine-learning-databases/statlog/german/german.data", 
              col_names = c("CheckingAccountStatus", 
                            "Duration", 
                            "CreditHistory", 
                            "Purpose", 
                            "Amount", 
                            "SavingsAccountBonds", 
                            "EmploymentDuration", 
                            "InstallmentRatePercentage", 
                            "Personal", 
                            "OtherDebtorsGuarantors", 
                            "ResidenceDuration", 
                            "Property", 
                            "Age", 
                            "OtherInstallmentPlans", 
                            "Housing", 
                            "NumberExistingCredits", 
                            "Job", 
                            "NumberPeopleMaintenance", 
                            "Telephone", 
                            "ForeignWorker", 
                            "Class"))
glimpse(german_credit)
```

