---
title: Dummy to Factor
author: Caroline Ledbetter
date: '2020-01-07'
slug: dummy-to-factor
categories:
  - R
tags:
  - tidyverse
lastmod: '2020-01-07T12:44:44-07:00'
description: 'Converting Dummy Variables to Factors'
show_in_homepage: yes
show_description: yes
license: ''
featured_image: ''
featured_image_preview: ''
comment: yes
math: no
---

```{css, echo = FALSE}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

**Motivation**  
I often have raw data where a variable has been provided as dummy variables 
as opposed to a character/factor variable. For example `job_employed = 1`, 
`job_unemployed = 0` instead of `job = employed`. While some modeling 
algorithms require the use of dummy variables, this can make it challenging 
to detect missing/multiple data and to create tables. For example in the German 
Credit Data[^1], if you want a count of the *purpose of the loan* field, 
you first
have to `sum` and then `pivot_longer` and it doesn't tell you if any 
observation
has no values true or multiple values true.

```{r, message = FALSE}
library(tidyverse)
data(GermanCredit, package = "caret")
GermanCredit %>% summarise_at(vars(starts_with('Purpose')), sum)
GermanCredit %>% 
  summarise_at(vars(starts_with('Purpose')), sum) %>% 
  pivot_longer(cols = everything(), 
               names_to = 'Purpose', 
               values_to = 'count', 
               names_pattern = '^[^.]+[.](.*)')
```

Now, what if you want to look at *checking account status* vs 
*housing status*. This 
is very difficult/tedious to do without converting to factors. 

```{r}
GermanCredit %>% 
  summarise(Lt0_and_Rent = sum(CheckingAccountStatus.lt.0 & Housing.Rent), 
            `0to200_and_Rent` = sum(CheckingAccountStatus.0.to.200 & 
                                      Housing.Rent))
# et cetera.

```

To that end I have created a function to convert dummy variables to factors. 
It requires that your variables have a seperator so that it can detect the 
main variable name from the level. 
My original function was quite slow and I appreciated the extremely valuable
help provided by 
[Christophe Dervieux](https://community.rstudio.com/u/cderv/summary) via 
this 
[Rstudio community post](https://community.rstudio.com/t/convert-dummy-variables-to-factors/48356)
.

```{r}
fct_dummy <- function(data, 
                      variables = tidyselect::everything(), 
                      sep = '.') {
  variables <- rlang::enquo(variables)
  # transform to long format the dummy columns
  tmp <- 
    tidyr::pivot_longer(data, 
                        cols = intersect(tidyselect::contains(sep),  
                                         !!variables),
                        names_to = c("groups", "levels"),
                        names_pattern = paste0("^([^'", sep, "]*)[", 
                                               sep, "](.*)"))
  
  # get the groups name for column selection after
  groups <- unique(tmp$groups)
  
  
  # keep only non dummy value and do not keep temp value col
  tmp <- dplyr::select(
    dplyr::filter(tmp, value == 1),
    -value)
  
  # function to return 'multiple' if more than 1 value is present
  ret_multiple <- function(x){
    if(length(x) > 1) return('multiple')
    return(x)
  }
  
  
  # tranform to wide format   
  tmp <- tidyr::pivot_wider(
    tmp,
    names_from = groups, 
    values_from = levels, 
    values_fn = list(levels = ret_multiple))
  
  
  # convert to factors the groups column
  dplyr::mutate_at(
    tmp,
    groups,
    ~ forcats::as_factor(.)
  )
}
```


Here is the original German Credit data and the converted tibble. 
```{r} 
glimpse(GermanCredit)
new_dat <- fct_dummy(GermanCredit)
glimpse(new_dat)

```


It is now straightforward to look at groupings of factors. 

```{r}
new_dat %>% 
  count(Housing, CheckingAccountStatus)
```

We can also see if we have multiples or missing data points
```{r}
GermanCredit[1, 13]
GermanCredit[1, 13] <- 1
```

```{r}
GermanCredit[1, 57] 
GermanCredit[1, 57] <- 0
```

```{r}
new_dat <- fct_dummy(GermanCredit)
new_dat %>% 
  count(Housing, CheckingAccountStatus)
```

You can also use the variables argument to only convert certain variables. 
```{r}
dat <- rename_all(GermanCredit, str_replace, 
                  pattern = '[.]', replacement = '_')
glimpse(fct_dummy(dat, variables = starts_with('P'), sep = '_'))

```


I have not yet decided what to do with this function, if I want to put it in 
one of my own packages or try and add it to another package. Please feel free 
to use it you find it useful. 


[^1]: This post uses the 
[German Credit data](http://archive.ics.uci.edu/ml/datasets/Statlog+(German+Credit+Data)) 
from the University of California Irving Machine Learning Repository. If you 
have the `caret` package installed, it is included. 
